\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-05 paper/draft.tex

To address this challenge, we develop \sys, a system that extracts data-parallelism from POSIX shell scripts.
% \kk{We need to mention data-parallelism earlier. We don't focus on general parallelization but rather data-parallelism (performing operations on disjoint partitions of the input and then merging the results. WE need to give a brief idea about that early on so that we don't create false ideas.}
% \nv{First attempt here. I am trying to get a copy out ASAP, we can refine this later}
% \kk{Mention that it outputs shell scripts.}
% \nv{I think the ``how'' it solves it should come later, we're still trying to convince them we solve it (see next sentences in this paragraph)
% \sys offers both classes of users in the second group such much-needed automation.
\sys benefits both programmer groups, % mentioned earlier, 
  with a particular emphasis on users.
Command developers are armed with a set of abstractions, akin to lightweight type annotations, for expressing the parallelizability properties of their commands:
  rather than providing a command's full observable behavior, these annotations focus primarily on their interaction with state.
Shell users, on the other hand, are provided with full automation:
\sys  analyzes their scripts and extracts latent parallelism.
% \sys extracts parallelism latent in their scripts by analyzing their parallelizability properties. % of their fragments.
% \kk{eurosys: I replaced the above to avoid saying that it analyzes parallelizability properties since it feels kinda redundant.}
% benefiting scripting experts and non-experts alike by offering automated POSIX-aware parallelization of shell-scripts as long as command developers
\sys's transformations are conservative, in that they do not attempt to parallelize fragments that lack sufficient information---\ie at worst, \sys will choose to not improve performance rather than risking breakage.
% \kk{Should we mention that we only show this empirically?}
% \nv{I don't think so; I don't think this is the right point to talk about our evaluation, we just state our intention/high-level principles.}

To address cold-start issues, \sys comes with a library of parallelizability semantics for commands in POSIX and GNU Coreutils.
% \kk{eurosys: Suggestion: library of annotations, annotation library, annotation thesaurus for \several/ commands in POSIX and GNU... I find the term parallelizability semantics confusing.}
These large classes of commands serve as the shell's standard library, expected to be used pervasively. % in scripts found out in the wild.
The study that led to their characterization also informed \sys's annotation and transformation components.\footnote{
  In fact, the scope of the study is broader, also incorporating commands outside POSIX and GNU Coreutils such as \ttt{curl} and \ttt{pandoc}.
}

These two components are finally paired with \sys's runtime component.
% \kk{eurosys: It is not totally clear what the phrase: "These two components" refers to. Maybe we can say that everything is tied together by \sys's runtime component?}
% \sys's runtime  component additional packs
% Aside from these two components, \sys packs a runtime component that serves dual purpose:
Aware of the \unix philosophy and abstractions, it packs a small library of highly-optimized data aggregators as well as high-performance primitives for eager data splitting and merging.
These address many practical challenges % such as task parallelism, deadlocks, and performance bottlenecks,
  and were developed by uncovering several pathological situations, on a few of which we report.

@@

We evaluate \sys using
  (i) a series of benchmarks, ranging from classic \unix one-liners to modern data-processing scripts;
  (ii) two large and complex use cases for temperature analysis and web indexing;
%% , and container orchestration;
and 
  (iii) a series of micro-benchmarks inspired by (and
comparing with) possible alternatives.  The results show
significant advantages in correctness, manual effort,
and performance---with an order-of-magnitude performance
improvement due to \sys's transformations and runtime primitives.
%% ---with \sys's transformations offering an order-of-magnitude performance improvements aided further by \sys's optimized runtime components.
Improvements depend on several factors such as the initial script execution time, the pipeline depth, and the parallelizability properties of its commands.
% \kk{eurosys: the pipeline depth seems to specific to pipelines. What we actually want to say is "how much parallelism did the initial script achieve (due to & and |). However I think that it is fine mentioning only two factors, "execution time, and the parallelizability properties of its commands."}
% \kk{eurosys: I removed parallelizability classes since we haven't introduced them until now.}

@@

  \item
  \S\ref{impl} details \sys's implementation, discussing several challenges related to \sys's translation, optimization, and runtime components and their solutions.
  % \kk{eurosys: suggestion: details \sys's runtime component, discussing several challenges related to correctness and performance as well as their solutions. The reason is that we don't talk about translation anymore and optimization is a bit too general.}
% \kk{Rewrite that when we know for sure what the section contains.}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{\detokenize{./figs/dish_overview.pdf}}
\caption{
  \textbf{\sys overview and correspondence to sections.}
  \sys identifies potentially \todo{parallelizable regions}~\sx{bg}, 
  converts them to a dataflow graph~\sx{graph-components}, applies iterative
  transformations~\sx{ir:transformations} based on the parallelizability properties (\S\ref{cmd}, \S\ref{ext}) 
  of the commands used in these regions, and finally emits a parallel script
  that uses custom primitives~\sx{optimizer}.
  % \kk{eurosys: Can we put GNU under POSIX so that the DFG and transformations move towards the center to become more prominent?Also we need to remember to edit the term parallelizable regions and the section in which it appears}
}
\vspace{-15pt}
\label{fig:overview}
\end{figure}

@@

At a high level, \sys takes as input a POSIX shell script like the one in Fig.~\ref{fig:example} and outputs a new script, a fragment of which is shown in Fig.~\ref{fig:example2}.
The new script may incorporate parallelism, and is handed off to the user's original shell interpreter % ---\eg \ttt{bash} or \ttt{zsh}---
  for execution.
% To expose and exploit data parallelism, % in a given shell script,
% \sys analyzes individual program fragments, addressing the challenges mentioned earlier.
% \kk{eurosys: I deleted this sentence since it doesn't offer a lot now that we have the overview below.}
An overview of \sys is shown in Fig.~\ref{fig:overview} and described below.
First, \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model~\sx{dataflow-regions} by identifying and transforming special regions termed \emph{parallelizable regions} and explained below. % nv idea: let's call them may-par regions
% \kk{eurosys: I suggest to replace the above sentence with the one below. There are a couple reasons. First I don't like the term source-to-source compiler because this sentence only describes the shell-graph compilation pass. Second, I would like to have some intuition about what there regions are (rather than deferring explanation for later). Third each region is tranformed to a DFG and not the whole script.}
% \sys's \todo{compiler} first identifies sections of the script that lack any synchronization or scheduling constraints and transforms them to a dataflow graph (DFG) model~sx{dataflow-regions}. 
It then performs a series of graph transformations that aim at exposing parallelism without breaking semantics~\sx{ir:transformations}.
% \kk{eurosys: I changed this sentence because I didn't like the term analyses, and because it wasn't clear that the transformations are performed on the graph.}
Finally, \sys converts these DFGs back to a shell script that it augments with its runtime primitives to improve performance~\sx{optimizer}.
% \kk{eurosys: I changed the last sentence because I don't like the term runtime library.}
% , and the technical challenges addressed to achieve performance benefits

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{\detokenize{./figs/pash_overview.pdf}}
\caption{
  \textbf{\sys overview and correspondence to sections.}
  \sys identifies potentially parallelizable regions~\sx{bg}, 
  converts them to a dataflow graph~\sx{graph-components}, applies iterative
  transformations~\sx{ir:transformations} based on the parallelizability properties (\S\ref{cmd}, \S\ref{ext}) 
  of the commands used in these regions, and finally emits a parallel script
  that uses custom primitives~\sx{optimizer}.
  % \kk{eurosys: Can we put GNU under POSIX so that the DFG and transformations move towards the center to become more prominent?Also we need to remember to edit the term parallelizable regions and the section in which it appears}
}
\vspace{-15pt}
\label{fig:overview}
\end{figure}

To address this challenge, we develop a system called \sys and outlined in Fig.~\ref{fig:overview} for parallelizing POSIX shell scripts.
% \kk{We need to mention data-parallelism earlier. We don't focus on general parallelization but rather data-parallelism (performing operations on disjoint partitions of the input and then merging the results. WE need to give a brief idea about that early on so that we don't create false ideas.}
% \nv{First attempt here. I am trying to get a copy out ASAP, we can refine this later}
% \kk{Mention that it outputs shell scripts.}
% \nv{I think the ``how'' it solves it should come later, we're still trying to convince them we solve it (see next sentences in this paragraph)
% \sys offers both classes of users in the second group such much-needed automation.
\sys benefits both programmer groups, % mentioned earlier, 
  with emphasis on users.
Command developers are given a set of abstractions, akin to lightweight type annotations, for expressing the parallelizability properties of their commands:
  rather than expressing a command's full observable behavior, these annotations focus primarily on its interaction with state.
Shell users, on the other hand, are provided with full automation:
  \sys  analyzes their scripts and extracts latent parallelism.
% \kk{eurosys: I replaced the above to avoid saying that it analyzes parallelizability properties since it feels kinda redundant.}
% benefiting scripting experts and non-experts alike by offering automated POSIX-aware parallelization of shell-scripts as long as command developers
\sys's transformations are conservative, in that they do not attempt to parallelize fragments that lack sufficient information---\ie at worst, \sys will choose to not improve performance rather than risking breakage.
% \kk{Should we mention that we only show this empirically?}
% \nv{I don't think so; I don't think this is the right point to talk about our evaluation, we just state our intention/high-level principles.}

To address cold-start issues, \sys comes with a library of parallelizability annotations for commands in POSIX and GNU Coreutils.
% \kk{eurosys: Suggestion: library of annotations, annotation library, annotation thesaurus for \several/ commands in POSIX and GNU... I find the term parallelizability semantics confusing.}
These large classes of commands serve as the shell's standard library, expected to be used pervasively. % in scripts found out in the wild.
The study that led to their characterization also informed \sys's annotation and transformation components.
% \footnote{
%   In fact, the scope of the study is broader, also incorporating commands outside POSIX and Coreutils such as \ttt{builtin}s, \ttt{curl} and \ttt{pandoc}.
% }

Everything is tied together with \sys's runtime component.
% \kk{eurosys: It is not totally clear what the phrase: "These two components" refers to. Maybe we can say that everything is tied together by \sys's runtime component?}
% \sys's runtime  component additional packs
% Aside from these two components, \sys packs a runtime component that serves dual purpose:
Aware of the \unix philosophy and abstractions, it packs a small library of highly-optimized data aggregators as well as high-performance primitives for eager data splitting and merging.
These address many practical challenges % such as task parallelism, deadlocks, and performance bottlenecks,
  and were developed by uncovering several pathological situations, on a few of which we report.

@@

We evaluate \sys using
  (i) a series of benchmarks, ranging from classic \unix one-liners to modern data-processing scripts;
  (ii) two large and complex use cases for temperature analysis and web indexing;
%% , and container orchestration;
and 
  (iii) a series of micro-benchmarks inspired by (and
comparing with) possible alternatives.  The results show
significant advantages in correctness, manual effort,
and performance---with 
speedups $X-X\times$ due to \sys's transformations and $Y-Y\times$ due to \sys's runtime primitives.
%% ---with \sys's transformations offering an order-of-magnitude performance improvements aided further by \sys's optimized runtime components.
Improvements depend on several factors such as a script's initial execution time, its pipeline depth, and the parallelizability properties of its commands.
% \kk{eurosys: the pipeline depth seems to specific to pipelines. What we actually want to say is "how much parallelism did the initial script achieve (due to & and |). However I think that it is fine mentioning only two factors, "execution time, and the parallelizability properties of its commands."}
% \kk{eurosys: I removed parallelizability classes since we haven't introduced them until now.}

@@

  \item
  \S\ref{impl} details \sys's library of runtime primitives, discussing the correctness and performance challenges it addresses.
  % \kk{eurosys: suggestion: details \sys's runtime component, discussing several challenges related to correctness and performance as well as their solutions. The reason is that we don't talk about translation anymore and optimization is a bit too general.}
% \kk{Rewrite that when we know for sure what the section contains.}

@@

@@

At a high level, \sys takes as input a POSIX shell script like the one in Fig.~\ref{fig:example} and outputs a new script, a fragment of which is shown in Fig.~\ref{fig:example2}.
The new script may incorporate parallelism, and is handed off to the user's original shell interpreter % ---\eg \ttt{bash} or \ttt{zsh}---
  for execution.
% concurrent version
% To expose and exploit data parallelism, % in a given shell script,
% %% parallel execution,
% \sys analyzes individual program fragments, addressing the challenges mentioned earlier.
% 
% \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model~\sx{ir} by identifying and transforming special regions termed \emph{parallelizable regions} explained below. % nv idea: let's call them may-par regions
% It then runs a series of iterative analyses and transformations within these regions aimed at pushing parallelism as far as possible without risking breaking semantics.
% Finally, the compiler converts the DFG back into a shell script, replacing nodes with instantiated commands or primitives taken from its runtime library~\sx{optimizer}.
To expose and exploit data parallelism, % in a given shell script,
  \sys analyzes individual program fragments, addressing the challenges mentioned earlier.
% \kk{eurosys: I deleted this sentence since it doesn't offer a lot now that we have the overview below.}
% An overview of \sys is shown in Fig.~\ref{fig:overview} and described below.
First, \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model~\sx{dataflow-regions} by identifying and transforming special regions termed \emph{parallelizable regions} and explained below. % nv idea: let's call them may-par regions
% \kk{eurosys: I suggest to replace the above sentence with the one below. There are a couple reasons. First I don't like the term source-to-source compiler because this sentence only describes the shell-graph compilation pass. Second, I would like to have some intuition about what there regions are (rather than deferring explanation for later). Third each region is tranformed to a DFG and not the whole script.}
% \sys's \todo{compiler} first identifies sections of the script that lack any synchronization or scheduling constraints and transforms them to a dataflow graph (DFG) model~sx{dataflow-regions}. 
It then performs a series of graph transformations that aim at exposing parallelism without breaking semantics~\sx{ir:transformations}.
% \kk{eurosys: I changed this sentence because I didn't like the term analyses, and because it wasn't clear that the transformations are performed on the graph.}
Finally, \sys converts these DFGs back to a shell script that it augments with its runtime primitives to improve performance~\sx{optimizer}.
% \kk{eurosys: I changed the last sentence because I don't like the term runtime library.}
% , and the technical challenges addressed to achieve performance benefits
